---
title: 'STAT 300: Written Assignment 2'
author: 'Aronn Grant Laurel (21232475)'
date: "April, 2025"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#if you do not have the package, type install.packages("name_of_the_package")
library(knitr)
library(zoo)
library(ggplot2)
```

### Question 1

(a) 

```{r, echo=TRUE}
student <- read.csv("students_data.csv", header = TRUE)
# head(student)

# Boxplot
ggplot(student, aes(x = TrainingMethod, y = ExamScore, fill = TrainingMethod)) +
  geom_boxplot() +
  labs(title = "Exam Score Distribution by Training Method",
       x = "Training Method",
       y = "Exam Score") +
  theme_minimal()

```
For our boxplot, In-Person Classes hold the highest median score and a relatively
wider inter-quartile range. Self-Study Method holdes the lowest median score while
Online Courses seem to have a smaller inter-quartile range. Overall,
students under the In-person class method seem to perform better on average.



```{r, echo=TRUE}
# Scatterplot
ggplot(student, aes(x = StudyHours, y = ExamScore, color = PreviousGPA)) +
  geom_point(alpha = 0.7) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Study Hours vs. Exam Score (Colored by Previous GPA)",
       x = "Study Hours",
       y = "Exam Score",
       color = "Previous GPA") +
  theme_minimal()
```
Overall, we can observe a slight positive correlation between exam scores
and study hours. Futhermore, we can see that higher GPA holders (3.5 - 4.0) tend to
cluster at the upper right of the graph with high exam score and study hours. 
On the other hand, lower GPA holders (3.0 - 0) tend to be at the bottom right
of the scatterplot whose students have lower exam scores and study hours.
Therefore, we can say that students who have higher exam scores.


(b)

```{r, echo=TRUE}
# this is where your R code goes
```


(c)

```{r, echo=TRUE}
# this is where your R code goes
```


### Question 2

(a) 

```{r, echo=TRUE}
# this is where your R code goes
```


(b)




(c)

```{r, echo=TRUE}
# this is where your R code goes
```

